---
title: "Instrumental Variables Examples"
date: "02-27-2025"
author:
- Prepared by Wendy Chan (Graduate School of Education, University of Pennsylvania)
- Answered by Huanxiao Wang
output:
  bookdown::pdf_document2:
    latex_engine: lualatex
  tufte::tufte_handout:
    latex_engine: xelatex
toc: true
toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

## Question 1

In the following exercises, we will use the \`\`SchoolingReturns" data from the \textsc{ivreg} package to estimate the effects of educational attainment on wages. We will use proximity to a four-year college (\textsc{nearcollege4}) as an instrumental variable.

```{r}
library(pacman)
p_load(ivreg)
data(SchoolingReturns)
```

Using \textsc{wage} as the outcome and \textsc{education} as the treatment, run the following analyses. (a) Estimate the effect of schooling on the \textbf{log} wage using OLS without additional covariates. What are the estimate and standard error? (b) Now, include \textsc{nearcollege4} as an IV and estimate the average treatment effect of schooling on log wages. Describe any differences between the estimates in (a) and (b).

```{r}
#OLS without covariates
m1 <- lm(log(wage) ~ education,data = SchoolingReturns)
summary(m1)

#IV without covariates
m2 <- ivreg(log(wage) ~ education|nearcollege4, data = SchoolingReturns)
# It's better to write m2 <- ivreg::ivreg(xxx), 
# it's more clear to let R recognize the which function from which package
summary(m2)
```

From the OLS model(m1), the estimate of the education coefficient is 0.052 with a standard error of 0.0029. This implies that with each additional year of education, the log of wages increases by 0.052.

From the IV model(m2), the estimate of the education coefficient is 0.19 with a standard error of 0.023. The estimate is larger but the standard error is also larger. The p-value for the IV estimate is slightly larger, but still statistically significant.

## Question 2

Using the same IV model in (b), now estimate the effect of schooling on education but include \textsc{nearcollege2} as another instrument. What are the estimates of the average treatment effect and standard error?

```{r}
m3 <- ivreg::ivreg(log(wage) ~ education|nearcollege4+nearcollege2, 
                   data = SchoolingReturns)
summary(m3)
```

When we include both nearcollege4 and nearcollege2, the estimate of the education coefficient is now 0.19 with a standard error of 0.023. The p-value is still statistically significant.

```{r}
m4 <- ivreg::ivreg(formula = log(wage) ~ feducation + meducation 
                   |education|nearcollege4 + nearcollege2 + 
                     feducation + meducation, data = SchoolingReturns)

summary(m4)
```

When including the two covariates, we see that the estimate of education is now larger at 0.25 with a standard error of 0.045 and the estimate is still statistically significant, albeit with a larger p-value. The estimates for father’s and mother’s education are both negative with similar magnitudes of −0.043 and 0.040, respectively. This implies that the parents’ education has a negative effect on the log of wages.

